---
title: "week1_R_solution.R"
author: "Nate George"
date: "January 9, 2018"
output: pdf_document
---

Load libraries we will use.
```{r load_libs, message=F}
library(data.table)
library(DMwR)
library(corrplot)
library(raster)
library(ggplot2)
```

Load the data; replace the placeholder -9 with NA.  Throw out the dm column (history of diabetes) because almost all values are missing.
```{r load_data}
fn <- '/home/nate/Dropbox/MSDS/MSDS680_ncg_S8W1_18/week1/heart.disease.data'
dt <- fread(fn)
summary(dt)
str(dt)
dt <- dt[, lapply(.SD, as.numeric)]
# replace all -9 with NA
dt[dt == -9] <- NA
summary(dt)
dim(dt)
# almost all 'dm' values are missing, so throw out that column
dt[, dm:=NULL]
```

Impute NA values with KNN.
```{r impute}
# impute missing values
dt.nona <- knnImputation(dt)
summary(dt.nona)
```

Examine the data with histograms.
```{r histograms}
# cholesterol and cigs appear to have large outliers
labels <- colnames(dt.nona)
for (i in seq(dim(dt)[2])) {
  col.data <- dt.nona[, get(labels[i])]
  nlevs <- nlevels(as.factor(col.data))
  if (nlevs <= 10) {
    barplot(table(col.data), main = NULL, xlab = labels[i])
    # axis(1, at=seq(nlevs), labels=levels(as.factor(col.data)))
  } else {
    hist(as.numeric(col.data), main = NULL, xlab = labels[i])
  }
  cat('\n\n')
  boxplot(col.data, main = labels[i])
  cat('\n\n')
}
```

Looking at a log-scale boxplot of all variables, we can see some low outliers on trestbps, chol, and thalach.
```{r boxplot}
ggplot(stack(dt.nona), aes(x = ind, y = values)) +
  geom_boxplot() + scale_y_continuous(trans = 'log10')
```

Outliers seem to be one low in thalach, a few high on cigs, and few high/low on cholestorol and trestbps.
Looking carefully, it seems there are some lower cigs outliers we don't want to remove, because these are probably accurate data.
We removed the upper ones because 100 cigarettes a day seems like an error in the data.
The assumption is the other outliers may be expreiment/data errors.


```{r outliers}
upper_outlier_cols <- c('trestbps', 'chol', 'cigs')
lower_outlier_cols <- c('thalach', 'chol', 'trestbps')
remove_outliers_up <- function(x) clamp(x, upper = boxplot.stats(x)$stats[5])
remove_outliers_low <- function(x) clamp(x, lower = boxplot.stats(x)$stats[1])

dt.nona[, upper_outlier_cols] <- dt.nona[, lapply(.SD, FUN = remove_outliers_up), .SDcols = upper_outlier_cols]
dt.nona[, lower_outlier_cols] <- dt.nona[, lapply(.SD, FUN = remove_outliers_low), .SDcols = lower_outlier_cols]

# you may have also noticed there is one weird 5.92... number in thal, this should be rounded to 6
dt.nona[, thal:=round(thal)]
# save data for later use
fn <- '~/Dropbox/MSDS/MSDS680_ncg_S8W1_18/week1/heart.disease.data.clean.csv'
fwrite(dt.nona, fn)
```

```{r boxplot2}
ggplot(stack(dt.nona), aes(x = ind, y = values)) +
  geom_boxplot() + scale_y_continuous(trans = 'log10')
```

```{r plots2}
for (i in seq(dim(dt)[2])) {
  col.data <- dt.nona[, get(labels[i])]
  nlevs <- nlevels(as.factor(col.data))
  if (nlevs <= 10) {
    barplot(table(col.data), main = NULL, xlab = labels[i])
    # axis(1, at=seq(nlevs), labels=levels(as.factor(col.data)))
  } else {
    hist(as.numeric(col.data), main = NULL, xlab = labels[i])
  }
  cat('\n\n')
  boxplot(col.data, main = labels[i])
  cat('\n\n')
}
```

We can see a number of things are positively correlated to the diagnosis, all except fasting blood sugar and cholesterol.
Heart rate is the only thing negatively correlated, so the higher the heart rate (during exercise) the less chance of heart disease.
```{r corrplot}
corrplot(cor(dt.nona), mar=c(3, 1, 1, 1))  # play around with margin parameters till it fits the screen
```
